On the application of microwave acoustic resonant viral inactivation \thanks{I would be delighted to include any criticisms or comments anyone may have; preferably leave them on the GitHub issues page, or email therobotist@gmail.com, @0xDBFB7 on Twitter, or irc.0xDBFB7.com:6667 \#covid.}

### 2021-02-11 documents/background.tex
Rotation

 (of course, it turns out )

@@

We put some

which establishes that - unique to certain viruses, and apparently unlike human cellular structures, as we shall see - coincidentally have just the right size, shape, stiffness, and net charge distribution to form a weak (Q=2) spherical dipole resonance mode which couples well to the microwave spectrum at approximately 8 GHz.

More critically, [] (and, in parallel, [Hung 2014]) theoretically model and then experimentally validate in various strains of Influenza A that - due to this acoustic-resonance effect - the power density levels required to crack the lipid envelope are near the safety limits for continuous exposure to humans.

There are a few minor issues with the experimental technique; sham, blinding, and dosimetry demanded by Vjl. are not mentioned.

\lettrine{U}nfortunately, we later realized that for both the absorption cross section and the requires the virus, a structure with not more than $6 \times 10^8$ atoms, to have an effective charge of $q=10^7$. It is very difficult to account for such a charge by any known mechanisms.

In another attempt to explain non-thermally, it is possible that this result could be explained by lipid membrane leakage, attributed to oxidation, which has been observed in strikingly similar regime in liposomes; but this would not necessarily explain the frequency-dependent inactivation seen in Figure 4b of \cite{Efficient2015}.

We are thoroughly mystified by the data presented. For the purposes of this paper, we accept a working hypothesis that this does not represent a clinically useful effect - especially as the exposures required are close to - or greatly exceed - the thermal limit for safety.

@@

One might be tempted to cite long wavelength of microwaves as an advantage. A UV photon is approximately the same wavelength as the virus itself;  In the case of, the wavelength is >> diameter, d

@@

If this mechanism exists, it would seem to provide significant advantages over existing UV or cold-plasma sterilization.

If the extensions made in our paper are valid, this is a non-ionizing, non-thermal

, non-chemical technique, harmless for continuous exposure to tissue, which can sterilize air and surfaces alike, including skin, eyes, and within hair; it evolves no ozone, can readily be generated with \$1 USD - scale devices, acts instantaneously, and - perhaps most critically - could be made to act {\bf within infected tissues}.

@@

So far as we are aware, there are no fundamental physical obstacles to non-thermal effects in biological systems. As an example, electroporation is a well-established - now a common laboratory procedure, in which an enormous electric field causes ions in solution to migrate across the membrane,

Such a non-thermal mechanism is quite unprecedented and extremely dubious, oft-posited but never substantiated; so it has taken some convincing that this is not a particularly abstruse artifact, especially when considered in the context of the broader non-thermal effects literature. For instance, the very paper (Edwards) which spurred development of the hydration-layer damping hypothesis used to explain these results, is has been quite\cite{Resonances1987} definitively\cite{Microwave1993a} shown to be an artifact, one which we unfortunately have not excluded in our study.\ In fact, even in light of the fine evidence that the other groups have collected, and the poor-quality data we have collected, we are still not entirely convinced that this effect exists - but perhaps this skepticism is unjustified.\\

\lettrine{It} cannot be overstated how unexpected and positively dubious this finding appears to be - that is, based on our limited research and inexperience to date.

For even the RF power limits set by standards organizations like \cite{ICNIRP2020} \cite{IEEE2006} appear to be based on the observation that no significant resonance modes exist in biological tissues due to damping forces from the surrounding tissue \cite{Vibrational2002}. As we shall see, this is grounded in solid {\it in vitro} (albeit more limited {\it in vivo}) evidence.

@@

Additionally, other sterilization techniques may produce superior results with less faffing about and should be evaluated in the same context. For instance, data on far UV [Buonanno 2017] indicates safety.\\

@@

>>>>

These reports attempt to establish that, ostensibly unlike human cellular structures, the studied species of viruses coincidentally about the right size, shape, stiffness, and net charge distribution to form a weak (Q=2) spherical dipole resonance mode which couples well to the microwave spectrum at approximately 8 GHz.

They theoretically model and then experimentally demonstrate in various strains of Influenza A that - supposedly due to this acoustic-resonance effect - the power density levels required to crack the lipid envelope are near the safety limits for continuous exposure to humans.

There are a few minor issues with the experimental technique

These very striking results have attracted some recent attention. \cite{Burkhartsmeier} have observed a resonance of the same nature (see below), and work is proceeding by their team.

Such a non-thermal mechanism is quite unprecedented and extremely dubious, oft-posited but never substantiated; so it has taken some convincing that this is not a particularly abstruse artifact, especially when considered in the context of the broader non-thermal effects literature.

For even the RF power limits set by standards organizations like \cite{ICNIRP2020} \cite{IEEE2006} appear to be based on the observation that any resonance modes which might exist in biological tissues are far too overdamped due to viscous from the surrounding solvent to be of any importance \cite{Vibrational2002}; this is grounded in very solid {\it in vitro}, {\it in vivo} and theoretical evidence.

On the other hand, under clinical and laboratory circumstances, a few well-established non-resonant non-thermal effects have already been put to practical use. An example, electroporation is now a common laboratory procedure, in which an enormous electric field causes ions in solution to migrate across a host cell's membrane. A variant, irreversible electroporation, has been used in the clinic for tumor ablation. In vitro there is evidence to suggest that sub-cellular effects take place under truly tremendous fields.

\lettrine{U}nfortunately, we later realized that for both the absorption cross section and the inactivation mechanism to align as suggested requires the virus, a structure with not more than $6 \times 10^8$ atoms, to have an effective charge of $q=10^7$. It is very difficult to account for such a charge.

On the other hand, Uzunoglu have produced a formulation of the theoretical analysis which requires a very plausible charge of only $q=10^4$.

In another attempt to explain non-thermally, it is possible that this result could be explained by a sort of lipid membrane leakage, attributed to oxidation, which has been observed in strikingly similar regime in liposomes; but this would not necessarily explain the frequency-dependent inactivation seen in Figure 4b of \cite{Efficient2015}.

We are thoroughly mystified by the data presented. For the purposes of this paper, we adopt a working hypothesis that this does not represent a clinically useful effect - especially as the exposures required are close to - or greatly exceed - the thermal limit for safety.

Were such a mechanism to exist, it would seem to provide significant advantages over existing UV or cold-plasma sterilization;

@@

@@

One might be tempted to cite long wavelength of microwaves as an advantage. A UV photon is approximately the same wavelength as the virus itself;  In the case of, the wavelength is >> diameter, meaning that

@@

@@

So far as we are aware, there are no fundamental physical obstacles to non-thermal effects in biological systems.

@@

Additionally, other sterilization techniques may produce superior results with less faffing about and should be evaluated in the same context. For instance, data on far UV\cite{Germicidal2017} indicates safety.\\

@@

; sham, blinding, and dosimetry demanded by Vjl. are not mentioned.

### 2021-02-11 documents/brillouin_pulse_math.tex
log colorbar
Rotation

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{muscle_pulse_8cm_minimize_test2}
	\caption{Second.}
\end{figure}

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{propagated_waveforms}
	\caption{}
\end{figure}

>>>>

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{muscle_pulse_8cm_minimize_test2}
	\caption{}
\end{figure}

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{propagated_waveforms_detail}
	\caption{\\
		Some input (left) and output (right) waves (without considering oscillator coupling) transmitted through 6 cm of muscle.\\
		X axis is time; Y axis is amplitude.
		(a) Gaussian monopulse with 16 ps FWHM.\\
		(b) Optimized 10 GHz sawtooth wave. Note the slight high-frequency ripple.\\ This has a frequency of approximately 1.5 GHz.\\
		(c) Prototypical precursor 10 GHz sine burst.\\
		(d) Continuous 10 GHz sine tone.
}
\end{figure}

\begin{figure}[H]
	%	\makebox[\textwidth][c]{
	\includesvg[width=\textwidth]{propagated_waveforms}
	\caption{The same four pulses plotted against depth (Y axis) and amplitude (Z axis)}
\end{figure}

### 2021-02-11 documents/experiment_1.tex
Oh hey the brilluoin just had a bug
Rotation

We postulate the following. When the virus is lysed, the resonator is destroyed. This should be reflected in the impedance spectrum.
We expect a non-reversible decrease in the resonance peak after the application of power. This should also appear in the time-domain pulse envelope; we expect a "notch" on the rising edge, with a slow taper upwards, which does not repeat on the next pulse.

@@

One could imagine it to be otherwise. For instance, the lysis that occurs after phage infection of bacteria does little harm to the cell; only the permeability of the membrane is damaged\cite{GROWTH}. Similarly, the results of the previous studies could perhaps have been obtained without necessarily destroying the virus' envelope.

>>>>

We postulate the following. When the virus is lysed, the resonator is destroyed. This should be reflected in the impedance spectrum.
We would expect a non-reversible decrease in the resonance peak after the application of power. This should also appear in the time-domain pulse envelope; we expect a "notch" on the rising edge, with a slow taper upwards, which does not repeat on the next pulse.

@@

One could imagine it to be otherwise. For instance, the lysis that occurs after phage infection of bacteria does little harm to the host cell; only the permeability of the membrane is damaged\cite{GROWTH}. Similarly, the results of the previous studies could perhaps have been obtained without necessarily destroying the virus' envelope.

### 2021-02-11 documents/lessons_learned.tex
Rotation

It is interesting to think about the meta-issues surrounding this project. Beta-glo assay, not a few kilometers away the food bank ran out of food and people were not able.

We have seen time and again that investment in science and technology - even basic research, so distant from the mechanics of society - can produce outsized effects on quality-of-life. Both NASA and CERN, neither of which do much related to people, produce a 7-to-1 and 3-to-1 return on investment.

@@

You could fit a pandemic through a good-sized breakthrough. The vaccines are an excellent example, really. Moreover you learn more to be more useful in the future. Phage display  -

@@

Is it the avenue of investigation? Perhaps a lack of peer input. Maybe there were too few people involved - now that science is "bigger", we need bigger groups to efficiently manage it?

Maybe if you don't have sufficient resources (knowledge, equipment, peers) to persue something efficiently, maybe it's not worth you pursuing it at all.

>>>>

It is interesting to think about the meta-issues surrounding this project. While an unusable \$600 Beta-Glo assay went flat, not a few kilometers away the food bank ran out of supplies and people were not able to feed their families. That is unjust.

On the other hand, we have seen time and again that investment in science and technology - even basic research, so distant from the mechanics of society - can produce outsized effects on quality-of-life. Both NASA and CERN, neither of which do much related to people, produce 7-to-1 and 3-to-1 returns on investment.

@@

You could fit a pandemic through a good-sized breakthrough. The vaccines are an excellent example. Moreover, on a personal level you learn new skills and techniques, to be more useful to others in the future.

@@

Is it the avenue of investigation? An arrogance in decision making? Perhaps a lack of peer input. Maybe there were too few people involved - now that science is "bigger", we need bigger groups to efficiently manage it?

Maybe if you don't have sufficient resources, equipment, knowledge, experience, or peers to persue something efficiently, or competently, and developing it slowly would not aid greatly in diffusing or disseminating knowledge, maybe it's not worth you pursuing it at all.

### 2021-02-11 documents/paper.tex
Text and prop improvements
Oh hey the brilluoin just had a bug
Rotation

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c|c| } 
		\hline
		col1 & col2 & col3 \\
		\hline
		& cell2 & cell3 \\ 
		& cell5 & cell6 \\ 
		& Negative & 5 \\ 
		\hline
	\end{tabular}

	\caption{Data from testing on bacteriophage. Data rounded to the nearest thousand photon counts. Not thoroughly characterized. Repeatability between test conditions   Background was no. 
	The flurometer printed a constant stream of the integration, all within the random error between capture windows $\sigma=0.5$ kcounts. }
\end{table}

@@

@@

@@

@@

%\printbibliography[]
%\printbibliography[title={All references}]

>>>>

@@

such a waveform may seem implausible but can probably be produced by e.g. an Auston switch feeding a pulse shaping line.

@@

\subfile{experiment_1}

\subfile{experiment_2}

@@

\subfile{acknowledgement}

@@

% LC_ALL=C tr -dc '\0-\177' <references.bib >newfile && mv newfile references.bib

%\printbibliography[title={General fluidics resources}]
